---
title: "Data_cleaning"
author: "April Ridlon"
date: "2/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Set up chunk above is a way to set the default arguments for your rmarkdown doc 

```{r, warning= FALSE, message = FALSE}
library(dplyr)
library(tidyr)
```

To supress the warnings, you can use the FALSE commands above

# Read in data 
[We are using dta from Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal.](df35b.304.2. https://knb.ecoinformatics.org/view/df35b.304.2)

```{r}
catch.original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method= "libcurl"),stringsAsFactors = F)
```
For windows machines you need an extra url bit at the beginning 
this uses a hyperlink to read in the data directly

head(catch.original)

Good to keep an original copy of the data in the code so you can look at the original without having to read it in again

# %>% Pipes
Can read these as "then", applies the function to a former function
The shortcut is command shift m
%>% 

r chunk shortcut is command option i

#Clean and reshape

Remove marginal sum and notes column using 'select()'
The first argument is always the dataframe
```{r}
catch_clean <- catch.original %>% 
  # select(Region, Year, Chinook, Coho,Pink, Chum)- or a more efficient way:
  select(-All, -notesRegCode) %>%
#to change the Chinook from character to number 
#mutate(Chinook = as.numeric(Chinook))
#got an error, which we resolve below 
 mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>% 
  #where the value is I, change to 1, otherwise keep the value 
 mutate(Chinook = as.numeric(Chinook))


head(catch_clean)
```

##QA
Quality Assurance section to check for errors, fix things
something isn't right in Chinook column and we need to figure it out since we couldn't use as numnmeric to convert


```{r}
i <- which(is.na(catch_clean$Chinook))
i
```

The above asks which row contains a character/caused the issue?

```{r}
catch.original[i,]
```
The above shows which row (which is this case was only one)
Assuming that this is a true error (you'd look at the data)

Reshape from wide to Tall format using 'pivot_longer'

```{r}
catch_long <- catch_clean  %>%
  pivot_longer(cols = -c(Region,Year), 
               names_to = "species",
               values_to = "catch")
head(catch_long)
```

Pivot wider example
```{r}
catch_wide <- catch_long %>%
  pivot_wider(names_from = Year,
              values_from = catch)

head(catch_wide)
```

This makes a wide table with Year as the column names and catch as the values

To return to the above chunk, what if we want to rename columns? Use rename.
And if we want to adjust the way the values show up (e.g. 5 actually means 5 thousand in this dataset). Use mutate. Note that mutate creates a new column, but it will overwrite an existing column if you use the dsame column name as an existing column
```{r}
catch_long <- catch_clean  %>%
  pivot_longer(cols = -c(Region,Year), 
               names_to = "species",
               values_to = "catch") %>% 
rename(catch_thousands = catch) %>% 
#to change to values in the catch column
mutate(catch = catch_thousands *1000) %>% 
#to go back to using only catch as the column name
select(-catch_thousands)  

head(catch_long)
```

#group_by and summarize
Calculate the mean of each region or Species, etc.
```{r}
mean_region <- catch_long %>%
  group_by(Region) %>%
summarise(catch_mean = mean(catch), num_obs = n())

head(mean_region)
```


Or you could create a table grouped by species 
```{r}
mean_species <- catch_long %>%
  group_by(species) %>%
summarise(catch_mean = mean(catch), num_obs = n())

head(mean_species)
```


In what year was the total standing catch the highest?
Group by year, summarise by the SUM of the catch, and arrange in reverse order from default to see the highest # first

```{r}
mean_total <- catch_long %>%
  group_by (Year) %>%
summarise(total_catch = sum(catch))%>%
arrange(-total_catch)

mean_total

```


What year has the highest annual Sockeye catch?
```{r}
annual_sockeye <- catch_long %>%
  filter( species =="Sockeye") %>%
group_by (Year) %>%
summarise(total_catch = sum(catch))%>%
arrange(-total_catch)

annual_sockeye

```

#Join to regions table
Read in a new csv, and select the columns you want to do the join with
```{r}
region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", stringsAsFactors = FALSE) %>%
  select(code, mgmtArea)

head(region_defs)
```

Now to do a left hand join with catch_long 

```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)
```


