---
title: "GgplotTutorialSNAPP"
author: "April Ridlon"
date: "2/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Ggplot Practice aka Data Wrangling


```{r}
library(leaflet)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(scales) # install.packages("scales")
library(patchwork) # install.packages("patchwork")
```

#Loading the Data
```{r}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"

esc <- tryCatch(
    read.csv("data/escapement.csv", stringsAsFactors = FALSE),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"), stringsAsFactors = FALSE)
        return(esc)
    }
)

head(esc)
```
The above gives an error message but the tryCatch function says if this is already loaded locally, read it, if not it downloads it from the url

#Data Processing 

```{r}
annual_esc <- esc %>% 
  #to separate the year, month and day, which are all in one column right now
  # sep = says separate by - which is how the data are entered originally
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>% 
  #check to see that the separation went well (hold command and click on annual_esc above)
  #the Year is character right now, so use mutate to change it to numeric
  mutate(Year = as.numeric(Year)) %>% 
  #to do a bunch of columns at once you could use this syntax: as.numeric(c(Year, Month, Day))
  #Now group by species, region, and year (at once) 
  group_by(Species, SASAP.Region, Year) %>% 
  #Since we want a total sum daily count, we use summarize as so
  summarize(escapement = sum(DailyCount)) %>% 
  #We now want to select only certain species, so we'll use filter to keep only some 
  # %in% function says is this specific value inside a set? 
  filter(Species %in% c("Chinook", "Sockeye", "Chum", "Coho", "Pink"))

head(annual_esc)

#check number of rows
nrow(annual_esc)
```

keyboard shorcut: Hold command and click on the name of the dataframe, and it will come up/show in the window (as shorcut instead of view(dataframe)

#Data Visulization with ggplot
First look at the database
```{r}
ggplot(annual_esc, aes(x = Species, y = escapement)) +
  geom_col()
```

Changing the colors of the bars


```{r}
ggplot(annual_esc, aes(x = Species, y = escapement)) +
  #the fill goes inside the geom_col bc you are telling it to fill the bars
  geom_col(fill= "blue")
```

aes can be read as "mapping" or "axis"- it is what you are mapping to the plot - a variable
if its not a variable (e.g. color), then it goes to the geom

To create a bar graph by region 
```{r}
ggplot(annual_esc, aes(x = Species, y = escapement, fill = SASAP.Region)) +
  geom_col()
```

Looking into the Kodiac region over a timeseries, using geom_line and some tidyverse functions 

```{r}
annual_esc %>% 
  filter(SASAP.Region == "Kodiak") %>% 
  # the == here is a conditional test "where region is equal to Kodiak"
  ggplot(aes(x = Year, y = escapement, color = Species)) + 
  #to get a timeseries, just plotting Year on the xaxis
  #to add color by species, you specify color = Species 
    geom_line() +
  #to add points to the line graph for more clarity   
    geom_point()

```

Background, themes

```{r}
annual_esc %>% 
  filter(SASAP.Region == "Kodiak") %>% 
  # the == here is a conditional test "where region is equal to Kodiak"
  ggplot(aes(x = Year, y = escapement, color = Species)) + 
  #to get a timeseries, just plotting Year on the xaxis
  #to add color by species, you specify color = Species 
    geom_line() +
  #to add points to the line graph for more clarity   
    geom_point() +
ggtitle("Salmon Escapement for the Kodiak") +
   ylab("Escapement") +
#to get rid of the grey grids in back
   theme_bw() +
#set the pre-set theme first, because if you center the plot and them run theme_bw it would overwrite the title adjustment
#to get the title centered, use the plot.title, then to move the legend to the bottom use legend.position    
  theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom")

```

#Setting GGplot Themes

Save a theme

```{r}
my_theme <- theme_bw() +
   theme(plot.title = element_text(hjust = 0.5), 
         legend.position = "bottom",
         #to remove the legend title
         legend.title = element_blank())

annual_esc %>% 
  filter(SASAP.Region == "Kodiak") %>% 
  ggplot(aes(x = Year, y = escapement, color = Species)) + 
    geom_line() +
    geom_point() +
ggtitle("Salmon Escapement for the Kodiak") +
   ylab("Escapement") +
  my_theme
```

#Using Facets
```{r}
ggplot(annual_esc, aes(x = Year, y = escapement, color = Species)) + 
    geom_line() +
    geom_point() +
    scale_y_continuous(labels = comma) +
    facet_wrap(~SASAP.Region, scales = "free_y", ncol = 2) +
    ylab("Escapement") +
    my_theme
```

Put two plots side by side 

```{r}
#uses the pkg patchwork (already loaded in)
library(patchwork)

# Plot 1
p1 <- ggplot(annual_esc, aes(x = Species, y = escapement)) +
  geom_col() + 
  my_theme

# Plo2
p2 <- ggplot(annual_esc, aes(x = Species, y = escapement)) +
  geom_col(fill = "blue") +
    my_theme

#Create two different plots as objects as above
#then add the objects (plot) together in a very simple syntax and it 
# puts the plots to gether in one figure
p1 + p2
```

#Writing Functions

Write a function that computes the percentage of a number: n*p/100
```{r}
percent_function <- function(n,p){
 n*p/100}
#to check if its running, call it and give it numbers 
percent_function(100,10)
#[1] 10
```

Make the ratio factor an argument so we can also use it to compute 1/1000 
```{r}
ratio_function <- function(n,p,r = 100){
 n*p/r
}
#r = 100 sets the default as 100, but naming it r also allows you the felxibility to define r as anything (1000, 2000, etc), as we do below 
ratio_function (134, 3, 1000)
```

On the same script write a second function to compute a2-b2
```{r}
ab_function <- function(a,b) {
  a^2 - b^2
}
ab_function (2, 3)
```
Modify your function to compute the square root: sqrt(a2-b2)

```{r}
ab_function <- function(a,b) {
  sqrt(a^2) - sqrt(b^2)
}
ab_function (2, 3)
```
