---
title: "GgplotTutorialSNAPP"
author: "April Ridlon"
date: "2/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Ggplot Practice aka Data Wrangling


```{r}
library(leaflet)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(scales) # install.packages("scales")
library(patchwork) # install.packages("patchwork")
```

#Loading the Data
```{r}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"

esc <- tryCatch(
    read.csv("data/escapement.csv", stringsAsFactors = FALSE),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"), stringsAsFactors = FALSE)
        return(esc)
    }
)

head(esc)
```
The above gives an error message but the tryCatch function says if this is already loaded locally, read it, if not it downloads it from the url

#Data Processing 

```{r}
annual_esc <- esc %>% 
  #to separate the year, month and day, which are all in one column right now
  # sep = says separate by - which is how the data are entered originally
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>% 
  #check to see that the separation went well (hold command and click on annual_esc above)
  #the Year is character right now, so use mutate to change it to numeric
  mutate(Year = as.numeric(Year)) %>% 
  #to do a bunch of columns at once you could use this syntax: as.numeric(c(Year, Month, Day))
  #Now group by species, region, and year (at once) 
  group_by(Species, SASAP.Region, Year) %>% 
  #Since we want a total sum daily count, we use summarize as so
  summarize(escapement = sum(DailyCount)) %>% 
  #We now want to select only certain species, so we'll use filter to keep only some 
  # %in% function says is this specific value inside a set? 
  filter(Species %in% c("Chinook", "Sockeye", "Chum", "Coho", "Pink"))

head(annual_esc)

#check number of rows
nrow(annual_esc)
```

keyboard shorcut: Hold command and click on the name of the dataframe, and it will come up/show in the window (as shorcut instead of view(dataframe)

#Data Visulization with ggplot
First look at the database
```{r}
ggplot(annual_esc, aes(x = Species, y = escapement)) +
  geom_col()
```

Changing the colors of the bars


```{r}
ggplot(annual_esc, aes(x = Species, y = escapement)) +
  #the fill goes inside the geom_col bc you are telling it to fill the bars
  geom_col(fill= "blue")
```


